---
title: 'ALM Fall 2019: Exercise 2'
output: html_document
---

# Purpose 

These exercises are designed to provide practice using R and interpreting results for data management, descriptive statistics, and three common bivariate inferential statistical tests.  

## Exercises 

Follow the steps below to clean the data in the **transgender_hc_ch2.csv** file, create a table of appropriate descriptive statistics, and conduct a few bivariate analyses. The data are from the 2014 Behavioral Risk Factor Surveillance Survey (BRFSS). Before you start working with the data, spend a few minutes looking through the BRFSS website to find and review the 2014 codebook. 

(1) Complete a prolog

```{r}
####################################
# Project name: 
# Project purpose: 
# Code author name: 
# Date last edited: 
# Location of data used: 
####################################
```

(2) Import and summarize the `transgender_hc_ch2.csv` 2014 BRFSS data file

```{r}

```

(3) Select participants who are transgender and select the following variables: `TRNSGNDR`, `X_AGE80`, `PHYSHLTH`, `HLTHPLN1`

```{r}

```

(4) Based on the BRFSS codebook, recode missing values and add category labels appropriately for all variables in your small data frame

```{r}

```

(5) Check data types and change any that need to be changed; many R commands do not work with integers, so change integers to numeric 


```{r}

```

(7) Use `tableone` to develop a table of appropriate descriptive statistics including all variables in the small data set 

```{r}

```

(8) Conduct the appropriate statistical test to determine if there is an association between transgender status (`TRNSGNDR`) and insurance status (`HLTHPLN1`), interpret your results 

```{r}

```


(9) Conduct the appropriate statistical test to compare the age of BRFSS participants to a hypothesized age of 50 years old, interpret your results 

```{r}


```


(10) Conduct the appropriate statistical test to determine if the age of BRFSS participants is the same across transgender status groups, interpret your results  

```{r}

```

## Want more? Try the hacker questions (optional)

(H1) Find the mean or median value of PHYSHLTH (whichever is most appropriate) and recode to create a new factor variable with values above the mean or median as "poor physical health" and values below the mean as "good physical health"ù

```{r}

```








