---
title: "week-2-workshop"
output: html_document
---

## Import and review the data for today

* Note that the file extension should be **.csv** for this code to work

```{r}
# import the data using read.csv
nhanes.2012 <- read.csv(file = "nhanes_2011_2012_ch3.csv")

# examine the data set 
summary(object = nhanes.2012)
```

## Data management topic: selecting cases and variables 

```{r}
# open the tidyverse for data management
library(package = "tidyverse")

# make a smaller data frame with four variables 
nhanes.2012.cleaned <- nhanes.2012 %>%
    select(AUQ300, AUQ310, RIAGENDR, RIDAGEYR)

# check the new data frame 
summary(object = nhanes.2012.cleaned)
```

## Data management topic: selecting cases and variables 

```{r}
# add code to keep values of AUQ300 that are Yes and No
nhanes.2012.cleaned <- nhanes.2012 %>%
    select(AUQ300, AUQ310, RIAGENDR, RIDAGEYR) %>%
    filter(AUQ300 <= 2)

# check the new data frame 
summary(object = nhanes.2012.cleaned)
```

## Data cleaning: Fix data types and add labels 

```{r}
# add to the smaller data frame with four variables 
# to keep values of AUQ300 that are Yes and No
nhanes.2012.cleaned <- nhanes.2012 %>%
    select(AUQ300, AUQ310, RIAGENDR, RIDAGEYR) %>%
    filter(AUQ300 <= 2) %>%
    mutate(AUQ300 = recode_factor(.x = AUQ300,
                                `1` = 'Yes',
                                `2` = 'No')) %>%
    mutate(AUQ310 = recode_factor(.x = AUQ310,
                                `1` = "1 to less than 100",
                                `2` = "100 to less than 1000",
                                `3` = "1000 to less than 10k",
                                `4` = "10k to less than 50k",
                                `5` = "50k or more",
                                `7` = "Don't know",
                                `9` = "Refused")) %>%
    mutate(RIAGENDR = recode_factor(.x = RIAGENDR,
                                  `1` = 'Male',
                                  `2` = 'Female')) %>%
    rename(gun.use = AUQ300) %>%
    rename(rounds.fired = AUQ310) %>%
    rename(sex = RIAGENDR) %>%
    rename(age = RIDAGEYR) 


# check the recoding
summary(object = nhanes.2012.cleaned)

```

## Choosing mean or median

```{r}
# examine age distribution  
nhanes.2012.cleaned %>%
    ggplot(aes(x = age)) + 
    geom_histogram()
```

## Find the mean, median, std dev, IQR in tidyverse 

```{r}
# descriptive statistics for continuous variables
nhanes.2012.cleaned %>%
    drop_na(age) %>%
    summarize(mean.age = mean(x = age),
              sd.age = sd(x = age),
              med.age = median(x = age),
              iqr.age = IQR(x = age))
```

## Find frequencies and percentages in base R

```{r}
# descriptive statistics for categorical variables 
table(nhanes.2012.cleaned$gun.use) 
prop.table(x = table(nhanes.2012.cleaned$gun.use))
```

## Find frequencies and percentages in tidyverse  

```{r}
# descriptive stats for categorical in tidyverse
nhanes.2012.cleaned %>%
  group_by(gun.use) %>%
  summarize(freq.gun.use = n()) %>%
  mutate(perc.gun.use = 100*(freq.gun.use / sum(freq.gun.use)))
```

## Using tableone to get all the descriptive stats

```{r}
# open tableone package
library(package = "tableone")
```

```{r}
# create and print table 
gun.use.table <- CreateTableOne(data = nhanes.2012.cleaned)
print(x = gun.use.table)
```

## Update the table with median instead of mean

```{r}
# create table 
gun.use.table <- CreateTableOne(data = nhanes.2012.cleaned)

# show all levels for categorical 
# specify non-normal age variable
print(x = gun.use.table, 
      showAllLevels = TRUE,
      nonnormal = 'age')
```

## Step 2: Chi-squared test statistic 

```{r}
# compute test statistic with CrossTable
chisq.test(x = nhanes.2012.cleaned$sex,
           y = nhanes.2012.cleaned$gun.use)
```


## Ok, but what is the relationship? 

```{r}
# gun use by sex percentages
prop.table(x = table(nhanes.2012.cleaned$sex,
                     nhanes.2012.cleaned$gun.use),
           margin = 1)
```

## Ok, but what is the relationship? (graph)

```{r}
# gun use by sex
nhanes.2012.cleaned %>%
    ggplot(aes(x = sex, fill = gun.use)) + 
    geom_bar(position = "dodge")
```

## Step 2: Calculate the test statistic

```{r}
# select the gun users from the clean data
nhanes.2012.gun.users <- nhanes.2012.cleaned %>%
    filter(gun.use == "Yes")

# conduct the t-test
t.test(x = nhanes.2012.gun.users$age, mu = 43.7)

```

## Step 2: Calculate the test statistic 

```{r}
# conduct the ANOVA
oneway.test(formula = age ~ rounds.fired,
            data = nhanes.2012.gun.users,
            var.equal = TRUE)
```

## Add some context 

```{r}
# means by group 
nhanes.2012.gun.users %>% 
    drop_na(age) %>% 
    drop_na(rounds.fired) %>%
    group_by(rounds.fired) %>%
    summarize(mean.age = mean(age),
              sd.age = sd(age))
```

## Add some context 

```{r}
# make a fancy graph
nhanes.2012.gun.users %>%
  drop_na(age) %>% 
  drop_na(rounds.fired) %>%
  ggplot(aes(y = age, x = rounds.fired)) + 
  geom_jitter(aes(color = rounds.fired), alpha = .6) +
  geom_boxplot(aes(fill = rounds.fired), alpha = .4) + 
  scale_fill_brewer(palette = "Spectral", guide = FALSE) +
  scale_color_brewer(palette = "Spectral", guide = FALSE) +
  theme_minimal() + 
  coord_flip() +
  labs(x = "Number of rounds fires", y = "Age in years")
```



