---
title: "R Markdown"
author: "Laura Brugger"
date: "10/8/2019"
output: ioslides_presentation
---

## Agenda

- Overview of Rmd
- Embedding Code Chunks
- Embedding Plots
- Formatting Text
- Knitting Documents

## Overview

- R Markdown is a file format for making documents with R
- An R Markdown document is written in markdown
- Markdown is an easy-to-write plain text format and contains chunks of embedded R code
- When you click the **knit** button, a document will be generated that includes both content and any embedded R code chunks
- this is what we've been using all semester for .rmd files! 

## Embedding Code Chunks

- you can embed an R code chunk like this: 

```{r}
summary(cars)
```

## Embedding Plots

- you can also embed plots like this (we use `echo = FALSE` to suppress the code): 

```{r, echo = FALSE}
plot(cars)
```

## Markdown for formatted text (1/3)

- .Rmd files use text written in `markdown`
- `markdown` is a set of conventions for formatting plain text
- You can use markdown to format
  + bold and italic
  + lists
  + headers & Section titles
  + hyperlinks
  + lots more! 
  
## Markdown for formatted text (2/3)

- **headers** - Place one or more hashtags at the start of a line that will be a header (or sub-header) 
  +  `# Main Heading` 
  + `## Smaller Heading`
  +  `### Even Smaller Heading`

- **italicized and bold text** - surround italized text with asterisks `*like this*` and surround bold text with two asterisks `**like this**`
  
## Markdown for formatted text (3/3)

- **lists** - group lines into bullet points that begin with asterisks. Leave a blank line before the first bullet: 

This is a list: 

- ` * item 1`
- ` * item 2`
- ` * item 3`

* **hyperlinks** - surround links with brackets and provide the link target in parentheses like this: `[Github](www.github.com)`

## Rendering

- to transform your .Rmd file to html, pdf, or word document, click the `Knit` icon
- select your desired format
- `rmarkdown` will duplicate your text in the new file format using the instructions you provided

# Let's try it!

## You Try: 

Reproduce the following paragraph in a new RMD file and then try knitting to each output: 

**Cuttlefish** are marine molluscs in the *cephalopod* family. This family also includes: 

* squid
* octopuses 
* nautiluses

Visit the [Cuttlefish Wiki](https://en.wikipedia.org/wiki/Cuttlefish) page to read more! 



