---
title: "r Markdown Demo"
author: "Laura Brugger"
date: "10/21/2019"
output:
  html_document: default
---

## Embedding Code Chunks

- you can embed an R code chunk like this: 

```{r}
summary(cars)
```

## Embedding Plots

- you can also embed plots like this (we add `echo = FALSE` to suppress the code): 

```{r, echo = FALSE}
plot(cars)
```


## RMD for Formatting

1. Headers 

- Place one or more hashtags at the start of a line that will be a header (or sub-header) 

# Largest
## A bit smaller
### Even smaller
#### You get the picture
##### This is as small as it gets
###### For real this time

2. Italicized and Bold Text
- surround italicized text with asterisks *like this*
- surround bold text with double asterisks **like this**


3. Lists 

- group lines into bullet points that begin with asterisks. Leave a blank line before the first bullet: 

This is a list: 

* item 1
* item 2
* item 3

4. Hyperlinks 

- surround links with brackets and provide the link target in parentheses like this: [Github](www.github.com)

## Rendering

- to transform your .Rmd file to `html`, `pdf`, or `word document`, click the "Knit" icon
- select your desired format
- `rmarkdown` will duplicate your text in the new file format using the instructions you provided

## You Try: 

Reproduce the following paragraph in a new RMD file and then try knitting to each output: 

**Cuttlefish** are marine molluscs in the *cephalopod* family. This family also includes: 

* squid
* octopuses 
* nautiluses

Visit the [Cuttlefish Wiki](https://en.wikipedia.org/wiki/Cuttlefish) page to read more! 