---
title: "M&M ANOVA"
output: html_document
---

# Trick or Treat: Using ANOVA to help your kids get the most candy

```{r echo = FALSE}
# enter your data
type.mm <- c("peanut", "caramel", "plain", "peanut", "caramel", "plain", "plain", "caramel", "peanut", "plain", "peanut", 
             "caramel", "peanut", "caramel", "plain", "caramel", "plain", "peanut", "caramel", "plain", "peanut", "caramel",
             "peanut", "plain")
number.mm <- c(6, 7, 16, 6, 7, 14, 13, 6, 8,15, 7, 7, 7, 7, 15, 7, 15, 7, 8, 16, 8, 7, 9, 15)

# make a data frame
mm.data <- data.frame(type.mm, number.mm)
```


```{r message=FALSE, warning=FALSE,include=FALSE}
# boxplot of m&ms by type
library(package = "tidyverse")
mm.data %>%
  ggplot(aes(x = type.mm, y = number.mm, fill = type.mm)) +
  geom_jitter() +
  geom_boxplot() +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  labs(x = "M & M type", y = "Number of pieces per package",
       subtitle = "") +
  theme_bw()

# descriptive statistics
library(package = "tableone")
mm.table <- CreateTableOne(data = mm.data,
                           strata = "type.mm",
                           vars = c("number.mm"))
print(mm.table)
```


```{r include=FALSE}
# density plot by groups to check normality
mm.data %>%
  ggplot(aes(x = number.mm)) + 
  geom_density(aes(fill = type.mm)) + 
  facet_wrap(~ type.mm) + 
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  theme_bw()
```

```{r echo = FALSE, results = FALSE, message=FALSE, warning=FALSE}
# use Levene in car package to check homogeneity of variances
library(package = "car")
leveneTest(y = number.mm ~ type.mm, data = mm.data)
```



```{r echo = FALSE, results = FALSE}
# ANOVA of mean mms by mm type
number.by.type <- oneway.test(formula = number.mm ~ type.mm,
                              data = mm.data,
                              var.equal = TRUE)
number.by.type
```


```{r echo = FALSE, results = FALSE}
# Bonferroni post-hoc test
bonf.mms <- pairwise.t.test(x = mm.data$number.mm,
                            g = mm.data$type.mm,
                            p.adj = "bonf")
bonf.mms
```



```{r echo = FALSE, results = FALSE}
# figure out the order of categories for planned comparisons
levels(x = mm.data$type.mm)
```

```{r echo = FALSE, results = FALSE}
# write a contrast to compare plain to combined peanut & caramel
contrast1 <- c(-1, -1, 2)
```

```{r echo = FALSE, results = FALSE}
# link the contrast to the candy type variable
contrasts(x = mm.data$type.mm) <- contrast1
```

```{r echo = FALSE, results = FALSE}
# re-run the model using aov() 
num.by.type.aov <- aov(formula = number.mm ~ type.mm, 
            data = mm.data)

# apply the contrasts to the anova object num.by.type.aov
# give the contrast a good descriptive name of "Plain vs. others"
plain.vs.other.contr <- summary.aov(object = num.by.type.aov,
                                 split = list(type.mm = list("Plain vs. others" = 1)))
plain.vs.other.contr
```

```{r echo = FALSE, results = FALSE}
# examine the means for the combined groups
# recode and compute the means for high school and college groups
# the labels should be in the order of the categories as you found
# with the levels code above
mm.data %>%
  mutate(type.mm = factor(x = type.mm, labels = c("peanut & caramel",
                                                "peanut & caramel", "plain"))) %>% 
  group_by(type.mm) %>%
  summarize(m.num.candies = mean(x = number.mm),
            sd.num.candies = sd(x = number.mm))
```

## Study results

We counted the number of M&Ms in 24 packages of fun-size M&Ms for Halloween. There were a mean of 7 (sd = .53) in the caramel M&Ms packages, 7.25 (sd = 1.04) in the peanut M&Ms packages, and 14.88 (sd = .99) in the plain M&Ms packages (Table 1; Figure 1)

Table 1. Mean number of M&Ms by type in 24 packages of fun-size M&Ms.

```{r echo = FALSE}
# descriptive statistics
library(package = "tableone")
mm.table <- CreateTableOne(data = mm.data,
                           strata = "type.mm",
                           vars = c("number.mm"))
print(mm.table)

# boxplot of m&ms by type
library(package = "tidyverse")
mm.data %>%
  ggplot(aes(x = type.mm, y = number.mm, fill = type.mm)) +
  geom_boxplot() +
  geom_jitter() +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  labs(x = "M & M type", y = "Number of pieces per package",
       subtitle = "Figure 1. M & Ms per package for 24 packages of fun-sized M & Ms (2019)") +
  theme_bw()
```

A one-way ANOVA comparing the mean number of M&Ms in fun-size packages for Halloween was statistically significant (F(2,21) = 205.56; p < .05) indicating that the mean number of candies per package is different across M&M types. Bonferroni post-hoc tests found that the mean number of M&Ms in plain type package was statistically significantly (p < .05) different from the mean number in the peanut and caramel types (see Table 1 for means). A planned comparison grouping caramel and peanut types (m = 7.13; sd = .81) and comparing this grouped mean to the mean number in plain also found a statistically significant (p < .05) difference in means. The data did not meet the normality assumption but did meet the equal variances assumption, so these results should not be generalized outside the sample. It appears that, to get the most candy trick-or-treating, kids should choose plain M&Ms when given a choice, but to be sure, a different analysis strategy might be useful.









