---
title: 'ALM Fall 2019: Exercise 2 Key'
output: html_document
---

# Purpose 

These exercises are designed to provide practice using R and interpreting results for data management, descriptive statistics, and three common bivariate inferential statistical tests.  

## Exercises 

Follow the steps below to use the BRFSS data in the **transgender_hc_ch2.csv** file to clean the data, create a table of appropriate descriptive statistics, and conduct a few bivariate analyses. Spend a few minutes looking through the BRFSS website before beginning and find the codebooks. 

(1) Complete a prolog

```{r}
####################################
# Project name: Exercise 2
# Project purpose: Practice coding and bivariate review
# Code author name: Jenine Harris
# Date last edited: 9/6/2019
# Location of data used: C:\Users\jenine\Box\teaching\Teaching\Fall2019\week-2-materials
####################################
```

(2) Import and summarize the `transgender_hc_ch2.csv` 2014 BRFSS data file

```{r}
# import the data 
transHC <- read.csv("transgender_hc_ch2.csv")

# examine the data
summary(object = transHC)
```

(3) Select participants who are transgender and select the following variables: `TRNSGNDR`, `X_AGE80`, `PHYSHLTH`, `HLTHPLN1`

```{r}
# open tidyverse
library(package = "tidyverse")

# select transgender and variables
transHC.clean <- transHC %>%
  filter(TRNSGNDR <= 3) %>%
  select(TRNSGNDR, X_AGE80, PHYSHLTH, HLTHPLN1)

# check the work
summary(object = transHC.clean)
```

(4) Based on the BRFSS codebook, recode missing values and add category labels appropriately for all variables in your small data frame

```{r}
# select transgender and variables
transHC.clean <- transHC %>%
  filter(TRNSGNDR <= 3) %>%
  select(TRNSGNDR, X_AGE80, PHYSHLTH, HLTHPLN1) %>%
  mutate(HLTHPLN1 = recode_factor(.x = HLTHPLN1,
                                  `1` = 'Yes',
                                  `2` = 'No',
                                  `7` = 'Don\'t know/not sure/missing',
                                  `9` = 'Refused')) %>%
  mutate(TRNSGNDR = recode_factor(.x = TRNSGNDR,
                                  `1` = 'Male to female',
                                  `2` = 'Female to male',
                                  `3` = 'Gender non-conforming')) %>%
  mutate(PHYSHLTH = na_if(PHYSHLTH, 77)) %>%
  mutate(PHYSHLTH = na_if(PHYSHLTH, 99)) %>%
  mutate(PHYSHLTH = recode(PHYSHLTH, `88` = 0L))

         
# check the work 
summary(object = transHC.clean)
```

(5) Check data types and change any that need to be changed; many R commands do not work with integers, so change integers to numeric 


```{r}
# check data types
class(x = transHC$X_AGE80)
class(x = transHC$PHYSHLTH)

# change data types to numeric
transHC.clean <- transHC %>%
  filter(TRNSGNDR <= 3) %>%
  select(TRNSGNDR, X_AGE80, PHYSHLTH, HLTHPLN1) %>%
  mutate(HLTHPLN1 = recode_factor(.x = HLTHPLN1,
                                  `1` = 'Yes',
                                  `2` = 'No',
                                  `7` = 'Don\'t know/not sure/missing',
                                  `9` = 'Refused')) %>%
  mutate(TRNSGNDR = recode_factor(.x = TRNSGNDR,
                                  `1` = 'Male to female',
                                  `2` = 'Female to male',
                                  `3` = 'Gender non-conforming')) %>%
  mutate(PHYSHLTH = na_if(PHYSHLTH, 77)) %>%
  mutate(PHYSHLTH = na_if(PHYSHLTH, 99)) %>%
  mutate(PHYSHLTH = recode(PHYSHLTH, `88` = 0L)) %>%
  mutate(PHYSHLTH = as.numeric(PHYSHLTH)) %>%
  mutate(X_AGE80 = as.numeric(X_AGE80))

# check data types
class(x = transHC.clean$X_AGE80)
class(x = transHC.clean$PHYSHLTH)


```

(7) Use `tableone` to develop a table of appropriate descriptive statistics including all variables in the small data set 

```{r}
# open tableone
library(package = "tableone")

# make and print table
transHC.table <- CreateTableOne(data = transHC.clean)
print(transHC.table)

```

(8) Conduct the appropriate statistical test to determine if there is an association between transgender status (`TRNSGNDR`) and insurance status (`HLTHPLN1`), interpret your results 

```{r}
# chi-squared 
chisq.test(x = transHC.clean$TRNSGNDR,
           y = transHC.clean$HLTHPLN1)

```

There is no statistically significant association between transgender status and health insurance status (chi-squared = 3.05; p = .80)

(9) Conduct the appropriate statistical test to compare the age of BRFSS participants to a hypothesized age of 50 years old, interpret your results 

```{r}
# one-sample t-test
t.test(x = transHC.clean$X_AGE80, mu = 50)
```

**The mean age in the sample was 53.6 years old. A one-sample t-test indicated that this was statistically significantly different from the hypothesized value of 50 years old (t = 5.6; p < .05).**

(10) Conduct the appropriate statistical test to determine if the age of BRFSS participants is the same across transgender status groups, interpret your results  

```{r}
# ANOVA
oneway.test(formula = X_AGE80 ~ TRNSGNDR,
            data = transHC.clean,
            var.equal = TRUE)
```

There is no statistically significant difference in mean age across groups of transgender status (F = 1.52; p = .22).

```{r}
# graph ages by transgender groups
transHC.clean %>%
  ggplot(aes(y = X_AGE80, x = TRNSGNDR)) + 
  geom_jitter(aes(color = TRNSGNDR), alpha = .6) +
  geom_boxplot(aes(fill = TRNSGNDR), alpha = .4) + 
  scale_fill_brewer(palette = "Spectral", guide = FALSE) +
  scale_color_brewer(palette = "Spectral", guide = FALSE) +
  theme_minimal() + 
  coord_flip() +
  labs(x = "Transgender group", y = "Age in years")
```


## Want more? Try the hacker questions (optional)

(H1) Find the mean or median value of PHYSHLTH (whichever is most appropriate) and recode to create a new factor variable with values above the mean or median as “poor physical health” and values below the mean as “good physical health”

```{r}
# check a histogram to see if normal
transHC.clean %>%
  ggplot(aes(x = PHYSHLTH)) +
  geom_histogram()

# not normally distributed use median
transHC.clean <- transHC %>%
  filter(TRNSGNDR <= 3) %>%
  select(TRNSGNDR, X_AGE80, PHYSHLTH, HLTHPLN1) %>%
  mutate(HLTHPLN1 = recode_factor(.x = HLTHPLN1,
                                  `1` = 'Yes',
                                  `2` = 'No',
                                  `7` = 'Don\'t know/not sure/missing',
                                  `9` = 'Refused')) %>%
  mutate(TRNSGNDR = recode_factor(.x = TRNSGNDR,
                                  `1` = 'Male to female',
                                  `2` = 'Female to male',
                                  `3` = 'Gender non-conforming')) %>%
  mutate(PHYSHLTH = na_if(PHYSHLTH, 77)) %>%
  mutate(PHYSHLTH = na_if(PHYSHLTH, 99)) %>%
  mutate(PHYSHLTH = recode(PHYSHLTH, `88` = 0L)) %>%
  mutate(PHYSHLTH = as.numeric(PHYSHLTH)) %>%
  mutate(X_AGE80 = as.numeric(X_AGE80)) %>% 
  mutate(health = cut(PHYSHLTH,
                      breaks = c(-Inf, median(PHYSHLTH, na.rm = TRUE), Inf),
                      labels = c("good physical health", "poor physical health")))

# look at data
summary(object = transHC.clean)

```








