---
title: "exercises-3"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

(3) Import the data in the `water_educ_2015_who_unesco_ch8.csv` file 

```{r}
# import data
water.educ <- read.csv(file = "water_educ_2015_who_unesco_ch8.csv")

# view it
summary(object = water.educ)

```

(4) Make a table of descriptive statistics for all the variables in the data except for country. There are multiple ways to do this. You could take a subset of variables in a smaller data frame and make a table of them all, or you could read the help documentation for `CreateTableOne` to learn how to choose variables for a table with the `vars = ` argument. Be sure to use appropriate statistics for each variable. 

```{r}
# open tidyverse
library(package = "tidyverse")

# get some histograms to see distributions of everything
water.educ %>% 
  ggplot(aes(x = med.age)) + geom_histogram()
water.educ %>% 
  ggplot(aes(x = perc.1dollar)) + geom_histogram()
water.educ %>% 
  ggplot(aes(x = perc.basic2015sani)) + geom_histogram()
water.educ %>% 
  ggplot(aes(x = perc.safe2015sani)) + geom_histogram()
water.educ %>% 
  ggplot(aes(x = perc.basic2015water)) + geom_histogram()
water.educ %>% 
  ggplot(aes(x = perc.safe2015water)) + geom_histogram()
water.educ %>% 
  ggplot(aes(x = perc.in.school)) + geom_histogram()
water.educ %>% 
  ggplot(aes(x = female.in.school)) + geom_histogram()
water.educ %>% 
  ggplot(aes(x = male.in.school)) + geom_histogram()

# use tableone to make a table of all variables
library(package = "tableone")
desc.table <- CreateTableOne(data = water.educ, 
                             vars = c("med.age", "perc.1dollar",
                                      "perc.basic2015sani", "perc.safe2015sani",
                                      "perc.basic2015water", "perc.safe2015water",
                                      "perc.in.school", "female.in.school", 
                                      "male.in.school"))
print(desc.table, nonnormal = c("med.age", "perc.1dollar",
                                "perc.basic2015sani", "perc.safe2015sani",
                                "perc.basic2015water", "perc.safe2015water",
                                "perc.in.school", "female.in.school",
                                "male.in.school"))
```

(4) Use a scatterplot to examine the relationship between `male.in.school` and `female.in.school` 

```{r}
# plot male and female in school
water.educ %>% 
  ggplot(aes(x = male.in.school, y = female.in.school)) + 
  geom_point()

```


(5) Use a scatterplot to examine the relationship between `male.in.school` and `perc.basic2015water` 

```{r}
# plot male and female in school
water.educ %>% 
  ggplot(aes(x = male.in.school, y = perc.basic2015water)) + 
  geom_point()

```

(6) Based on the scatterplots from questions 4 and 5, make predictions about what you would find when you conduct Pearson correlation analyses for `male.in.school` and `female.in.school` and for `male.in.school` and `perc.basic2015water`.

(7) Conduct a Pearson's r correlation analysis for each pair of variables. Interpret each r statistic in terms of direction, size, and significance. 

```{r}
# correlations
cor.test(x = water.educ$male.in.school, 
         y = water.educ$female.in.school)
cor.test(x = water.educ$male.in.school, 
         y = water.educ$perc.basic2015water)
cor.test(x = water.educ$female.in.school, 
         y = water.educ$perc.basic2015water)

```

(8) Check assumptions for the Pearson's r for each pair of variables. 

```{r}
# males in school and females in school 
water.educ %>% 
  ggplot(aes(x = male.in.school, y = female.in.school)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE,
              aes(color = "Fit line")) +
  geom_smooth(method = "loess", se = FALSE,
              aes(color = "Loess curve")) +
  theme_minimal() +
  scale_color_manual(values = c("gray60", "deeppink"),
                     name = "")

# males in school and basic water access
water.educ %>% 
  ggplot(aes(x = male.in.school, y = perc.basic2015water)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE,
              aes(color = "Fit line")) +
  geom_smooth(method = "loess", se = FALSE,
              aes(color = "Loess curve")) +
  theme_minimal() +
  scale_color_manual(values = c("gray60", "deeppink"),
                     name = "")

# females in school and basic water access 
water.educ %>% 
  ggplot(aes(x = female.in.school, y = perc.basic2015water)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE,
              aes(color = "Fit line")) +
  geom_smooth(method = "loess", se = FALSE,
              aes(color = "Loess curve")) +
  theme_minimal() +
  scale_color_manual(values = c("gray60", "deeppink"),
                     name = "")

```

(9) If assumptions are not met for the Pearson's r, conduct and interpret a Spearman's correlation analysis including assumption checking. Interpret your results. 

```{r}
# correlations
cor.test(x = water.educ$male.in.school, 
         y = water.educ$female.in.school,
         method = "spearman")
cor.test(x = water.educ$male.in.school, 
         y = water.educ$perc.basic2015water,
         method = "spearman")
cor.test(x = water.educ$female.in.school, 
         y = water.educ$perc.basic2015water,
         method = "spearman")
```





