---
title: "Exercise 5"
output: html_document
---

# Purpose 

These exercises are designed to practice conducting and interpreting a multiple linear regression model. 

## Exercises 

* Save this file and the `dist_ssp_amfar_ch9.csv` data file 
* If you have not already, read the readings for this week

(1) Based on your readings for this week, what are dummy variables and how are they used in linear regression? Cite the reading(s) you used to answer this question.

(2) Copy the code for the simple linear regression model you build in Exercise 4 and paste it into this document.

```{r}

```


(3) Starting with the model from Exercise 4, add the metro variable and the percent uninsured variable to the model. Run the model and write a full interpretation of what you found.


```{r}

```


(4) Use the partial-F test to compare the simple linear model from Exercise 4 (pasted above in #2) with the larger model from #3. Interpret the results. 


```{r}

```

