---
title: 'Applied Linear Modeling'
date: 'September 24, 2019'
---

```{r}
# distance to syringe program data
dist.ssp <- read.csv(file = "dist_ssp_amfar_ch9.csv")

# summary
summary(object = dist.ssp)
```

## Codebook 

Based on the amFAR website, the variables have the following meanings:

* county: the county name 
* STATEABBREVIATION: the two-letter abbreviation for the state the county is in 
* dist_SSP: distance in miles to the nearest syringe services program 
* HIVprevalence: people age 13 and older living with diagnosed HIV per 100,000 
* opioid_RxRate: number of opioid prescriptions per 100 people 
* pctunins: percentage of the civilian noninstitutionalized population with no health insurance coverage 
* metro: county is non-metro, which includes open countryside, rural towns, or smaller cities with up to 49,999 people, or metro 

## A research question 

**How can uninsurance, metro or non-metro status, HIV prevalence, and number of opioid prescriptions predict or explain distance to the nearest syringe program at the county level?**

## But first, descriptives 

```{r}
# make a table of descriptives
library(package = "tableone")

# dist_SSP and HIVprevalence are skewed
ssp.table <- CreateTableOne(data = dist.ssp,
                         vars = c('dist_SSP', 'HIVprevalence',
                                  'opioid_RxRate', 'pctunins',
                                  'metro'))
print(ssp.table, nonnormal = c("dist_SSP", "HIVprevalence"),
      showAllLevels = TRUE)
```

## Revisiting our original model  

```{r}
# linear regression of distance to syringe program by percent uninsured
dist.by.unins <- lm(formula = dist_SSP ~ pctunins, 
                data = dist.ssp, na.action = na.exclude)
summary(dist.by.unins) 
```

## Revisiting the confidence intervals 

```{r}
# confidence interval for regression parameters
ci.dist.by.unins <- confint(dist.by.unins)
ci.dist.by.unins
```

## Adding a binary variable to the model 

```{r eval = FALSE} 
# open the tidyverse
library(package = "tidyverse")

# metro and distance to SSP
dist.ssp %>%
  ggplot(aes(x = metro, y = dist_SSP, fill = metro)) +
  geom_jitter(aes(color = metro), alpha = .6) +
  geom_boxplot(aes(fill = metro), alpha = .4) + 
  labs(x = "Type of county", 
       y = "Miles to syringe program") +
  scale_fill_manual(values = c("#78A678", "#7463AC"), guide = FALSE) + 
  scale_color_manual(values = c("#78A678", "#7463AC"), guide = FALSE) +
  theme_minimal() 
```

## Distance to syringe program and metro status 

```{r echo = FALSE} 
library(package = "tidyverse")
# metro and distance to SSP
dist.ssp %>%
  ggplot(aes(x = metro, y = dist_SSP, fill = metro)) +
  geom_jitter(aes(color = metro), alpha = .6) +
  geom_boxplot(aes(fill = metro), alpha = .4) + 
  labs(x = "Type of county", 
       y = "Miles to syringe program") +
  scale_fill_manual(values = c("#78A678", "#7463AC"), guide = FALSE) + 
  scale_color_manual(values = c("#78A678", "#7463AC"), guide = FALSE) +
  theme_minimal() 
```

## Adding metro to the model 

```{r}
# linear regression distance to syringe program by
# uninsured percent and metro status in 500 counties
dist.by.unins.metro <- lm(formula = dist_SSP ~ pctunins +
                        metro, data = dist.ssp)
summary(object = dist.by.unins.metro)
```

## Get some confidence intervals 

```{r}
# confidence interval for regression parameters
ci.dist.by.unins.met <- confint(dist.by.unins.metro)
ci.dist.by.unins.met
```

## Visualizing the model 

```{r}
# graphing the regression model with percent uninsured and metro
dist.ssp %>%
  ggplot(aes(x = pctunins, y = dist_SSP, group = metro)) +
  geom_line(data = broom::augment(dist.by.unins.metro), 
            aes(y = .fitted, linetype = metro)) +
  geom_point(aes(color = metro), alpha = .4, size = 2) + 
  theme_minimal() + 
  scale_color_manual(values = c("dodgerblue2", "deeppink"), name = "County") +
  labs(y = "Miles to nearest syringe program", x = "County percent uninsured") +
  scale_linetype_manual(values = c(1,2), name = "Regression line\n(predicted values)") 
  
```

```{r warning=FALSE, message=FALSE}
# check normality of distance variable
dist.plot <- dist.ssp %>%
  ggplot(aes(x = dist_SSP)) + 
  geom_histogram(fill = "#7463AC", color = "white") + 
  theme_minimal() + 
  labs(x = "Miles to syringe program",
       y = "Number of counties",
       subtitle = "Distance to syringe program") 
# check normality of uninsured variable
perc.plot <- dist.ssp %>%
  ggplot(aes(x = HIVprevalence)) + 
  geom_histogram(fill = "#7463AC", color = "white") + 
  theme_minimal() + 
  labs(x = "HIV cases per 100,000",
       y = "Number of counties",
       subtitle = "HIV prevalence")

gridExtra::grid.arrange(dist.plot, perc.plot)

```

## Another way to check normality: Q-Q plots

```{r}
# Q-Q plot of distance variable to check normality
dist.ssp %>%
  ggplot(aes(sample = dist_SSP)) + 
  stat_qq(aes(color = "County"), alpha = .6) + 
  geom_abline(aes(intercept = mean(dist_SSP), slope = sd(dist_SSP), 
                  linetype = "Normally distributed"), color = "gray60", size = 1) +
  theme_minimal() + 
  labs(x = "Theoretical normal distribution",
       y = "Observed miles to syringe program") +
  scale_color_manual(values = "#7463AC", name = "") +
  scale_linetype_manual(values = 1, name = "")
```

## Another way to check normality: Q-Q plots

```{r}
# Q-Q plot of distance variable to check normality
dist.ssp %>%
  ggplot(aes(sample = HIVprevalence)) + 
  stat_qq(aes(color = "County"), alpha = .6) + 
  geom_abline(aes(intercept = mean(HIVprevalence), slope = sd(HIVprevalence), 
                  linetype = "Normally distributed"), color = "gray60", size = 1) +
  theme_minimal() + 
  labs(x = "Theoretical normal distribution",
       y = "HIV cases per 100,000") +
  scale_color_manual(values = "#7463AC", name = "") +
  scale_linetype_manual(values = 1, name = "")
```

## Square root transformation of distance  

```{r warning = FALSE, message=FALSE}
dist.ssp %>%
  ggplot(aes(x = sqrt(x = dist_SSP))) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Square root of distance", y = "")+
  theme_minimal()
```

## Cube root transformation of distance 

```{r warning = FALSE, message=FALSE}
 dist.ssp %>%
  ggplot(aes(x = (dist_SSP)^(1/3))) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Cube root of distance", y = "Number of counties") +
  theme_minimal()
```

## Inverse of distance 

```{r warning = FALSE, message=FALSE}
 dist.ssp %>%
  ggplot(aes(x = 1/dist_SSP)) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Inverse distance", y = "Number of counties") +
  theme_minimal()
```

## Log of distance 

```{r warning = FALSE, message=FALSE}
 dist.ssp %>%
  ggplot(aes(x = log(x = dist_SSP))) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Log of distance", y = "Number of counties") +
  theme_minimal()
```

## All transformations for distance 

```{r warning=FALSE, error=FALSE, message=FALSE}
# histogram of square root of dist_SSP
cube.root.dist <- dist.ssp %>%
  ggplot(aes(x = (dist_SSP)^(1/3))) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Cube root of distance", y = "Number of counties") +
  theme_minimal()
sq.root.dist <- dist.ssp %>%
  ggplot(aes(x = sqrt(x = dist_SSP))) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Square root of distance", y = "")+
  theme_minimal()
inverse.dist <- dist.ssp %>%
  ggplot(aes(x = 1/dist_SSP)) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Inverse of distance", y = "Number of counties")+
  theme_minimal()
log.dist <- dist.ssp %>%
  ggplot(aes(x = log(x = dist_SSP))) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Log of distance", y = "")+
  theme_minimal()

# this requires the gridExtra package to be installed
# view options for transformation
gridExtra::grid.arrange(cube.root.dist, sq.root.dist,
                        inverse.dist, log.dist)

```

## All transformations for HIV prevalence 

```{r warning = FALSE, message = FALSE}
# histogram of transformed HIVprevalence
# cube root
cube.root.dist <- dist.ssp %>%
  ggplot(aes(x = (HIVprevalence)^(1/3))) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Cube root of HIV prevalence", y = "Number of counties")+
  theme_minimal()
# square root
sq.root.dist <- dist.ssp %>%
  ggplot(aes(x = sqrt(x = HIVprevalence))) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Square root of HIV prevalence", y = "")+
  theme_minimal()
# inverse
inverse.dist <- dist.ssp %>%
  ggplot(aes(x = 1/HIVprevalence)) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Inverse of HIV prevalence", y = "Number of counties")+
  theme_minimal()
# log
log.dist <- dist.ssp %>%
  ggplot(aes(x = log(x = HIVprevalence))) +
  geom_histogram(fill = "#7463AC", col = "white") +
  labs(x = "Log of HIV prevalence", y = "")+
  theme_minimal()

# view options for transformation
gridExtra::grid.arrange(cube.root.dist, sq.root.dist,
                        inverse.dist, log.dist)

```

## Model with transformed variables 

```{r message = FALSE, warning = FALSE}
# linear regression of distance by percent uninsured, HIV prevalence,
# metro status
dist.full.model <- lm(formula = (dist_SSP)^(1/3) ~ pctunins +
                      log(x = HIVprevalence) + metro,
                      data = dist.ssp,
                      na.action = na.exclude)
summary(object = dist.full.model) 
```

## Making the models comparable 

```{r}
# drop observations with missing HIV data
dist.ssp.small <- dist.ssp %>%
  drop_na(HIVprevalence) %>%
  mutate(log.HIV = log(x = HIVprevalence)) %>%
  drop_na(log.HIV)
```
## Re-run the models with the new data frame 

```{r message = FALSE, warning = FALSE}
# re-run regressions with the smaller data frame
dist.unins.metro.trans <- lm(formula = (dist_SSP)^(1/3) ~ pctunins + metro,
                      data = dist.ssp.small)

dist.full.model.small <- lm(formula = (dist_SSP)^(1/3) ~ pctunins + 
                              metro + log(x = HIVprevalence),
                      data = dist.ssp.small)
```
    

## Conducting a Partial-F test

```{r}
# partial F test for dist.by.unins.metro and dist.full.model
anova(object = dist.unins.metro.trans, dist.full.model.small)
```

## A cool plot 

```{r}
# graphing the regression model with three predictors
dist.ssp %>%
  ggplot(aes(x = pctunins, y = dist_SSP, group = metro, size = HIVprevalence)) +
  geom_point(aes(color = metro), alpha = .4) + 
  theme_minimal() + 
  scale_color_manual(values = c("dodgerblue2", "deeppink"), name = "County") +
  scale_size(range = c(1, 10), name="HIV cases\nper 100,000") +
  labs(y = "Miles to nearest syringe program", x = "County percent uninsured")
```

