---
title: 'ALM week 9: logistic regression'
---

## Import the data

```{r echo=TRUE}
# import the libraries data file
libraries <- read.csv("pew_libraries_2016_cleaned_ch10.csv")

# check the data
summary(object = libraries)
```

## Codebook 

The data are publicly available from the Pew Internet & American Life website. The variables are: 

* **age:** age in years
* **sex:** sex (female/male)
* **rurality:** lives in rural/suburban/urban area
* **disabled:** has a disability (yes/no)
* **uses.lib:** has used library in the last year (yes/no) 
* **ses:** socioeconomic status (high/medium/low)
* **raceth:** race and ethnicity (Hispanic, Non-Hispanic Black, Non-Hispanic White)
* **educ:** highest education completed (< HS, HS to 2-year degree, Four-year degree or more)
* **parent:** parent status (parent/not parent)

## Exploratory data analysis: examine distributions 

```{r message = FALSE, warning = FALSE}
# open tidyverse
library(package = "tidyverse")

# examine the distribution of age 
libraries %>% 
  ggplot(aes(x = age)) + 
  geom_density(fill = "#7463AC", alpha = .6) + 
  theme_minimal() +
  labs(y = "Probability density", x = "Age in years")
```

## Exploratory data analysis: make a table 

```{r}
# open tableone
library(package = "tableone")
# create the table
table.desc <- CreateTableOne(data = libraries, 
                             strata = 'uses.lib',
                             vars = c("age", "sex", "parent", "disabled",
                                      "ses", "raceth", "educ", "rurality"))
print(table.desc, 
      nonnormal = 'age', 
      showAllLevels = TRUE)

```

## Clean the outcome for the regression model  

```{r}
# recode so yes = 1 and no = 0
libraries.cleaned <- libraries %>%
  mutate(uses.lib.num = recode(uses.lib,
                               `yes` = 1,
                               `no` = 0))

# check recoding
table(libraries.cleaned$uses.lib.num, libraries.cleaned$uses.lib)
```

## Estimate the model and get GVIF

```{r warning=FALSE, message=FALSE}
# predict library use 
libUseModel <- glm(uses.lib.num ~ age + sex + parent +
                     disabled + ses + raceth + educ + rurality,
                   data = libraries.cleaned,
                   family = "binomial")

# compute GVIF for libUseModel
library(car)
vif(mod = libUseModel)
```

## Check assumption 3: Linearity 

```{r message = FALSE, warning = FALSE}
# make a variable of the log odds of the predicted probabilities
logodds.use <- log(x = libUseModel$fitted.values/(1-libUseModel$fitted.values))

# make a small data frame with the log odds variable and the age predictor
linearity.data <- data.frame(logodds.use, age = libUseModel$model$age)

# create a plot with linear and actual relationships shown
linearity.data %>%
  ggplot(aes(x = age, y = logodds.use))+
  geom_point(aes(size = "Observation"), color = "gray60", alpha = .6) +
  geom_smooth(se = FALSE, aes(color = "Loess curve")) + 
  geom_smooth(method = lm, se = FALSE, aes(color = "linear")) + 
  theme_minimal() +
  labs(x = "Age in years", y = "Log-odds of library use predicted probability") +
  scale_color_manual(name="Type of fit line", values=c("dodgerblue2", "deeppink")) +
  scale_size_manual(values = 1.5, name = "")

```



## Model diagnostics 

```{r}
# use same code as from linear regression 
# change the cutoff for leverage to reflect the 13 parameters in the model 
libraries.cleaned.diag <- libraries.cleaned %>%
  drop_na() %>%
  mutate(standardres = rstandard(model = libUseModel)) %>%
  mutate(cooks.dist = cooks.distance(model = libUseModel)) %>%
  mutate(lever = hatvalues(model = libUseModel)) %>%
  mutate(outlier.infl = as.numeric(x = lever > 2*13/n()) +
           as.numeric(x = cooks.dist > 4/n()) +
           as.numeric(x = abs(x = standardres) > 1.96))

# examine the outliers & influential 
libraries.cleaned.diag %>%
  filter(outlier.infl >= 2)
```

    
## Model results 

```{r warning=FALSE, message = FALSE}
# use odds.n.ends to get model results
library(package = "odds.n.ends")
odds.n.ends(libUseModel)
```


    
## Model significance 


```{r echo = TRUE}
# get baseline probability
prop.table(x = table(libraries.cleaned$uses.lib))
```

















