---
title: 'Week 3 Workshop'
---

```{r}
# import the water data
water.educ <- read.csv(file = "water_educ_2015_who_unesco_ch8.csv")

# examine the data
summary(object = water.educ)
```

## Codebook 

* country: the name of the country 
* med.age: the median age of the citizens in the country 
* perc.1dollar: percentage of citizens living on $1 per day or less 
* perc.basic2015sani: percentage of citizens with basic sanitation access 
* perc.safe2015sani: percentage of citizens with safe sanitation access 
* perc.basic2015water: percentage of citizens with basic water access 
* perc.safe2015water: percentage of citizens with safe water access 
* perc.in.school: percentage of school-age people in primary and secondary school  
* female.in.school: percentage of female school-age people in primary and secondary school 
* male.in.school: percentage of male school-age people in primary and secondary school 

## Descriptive statistics 

```{r}
# open the tidyverse
library(package = "tidyverse")

# descriptive statistics for females in school and water access
water.educ %>%
  drop_na(female.in.school) %>%
  drop_na(perc.basic2015water) %>%
  summarize(m.f.educ = mean(x = female.in.school),
            sd.f.educ = sd(x = female.in.school),
            m.bas.water = mean(x = perc.basic2015water),
            sd.bas.water = sd(x = perc.basic2015water))
```

## Visualize the relationship 

```{r}
# explore plot of female education and water access
water.educ %>%
  ggplot(aes(y = female.in.school, x = perc.basic2015water)) + 
  geom_point(size = 2, color = "#7463AC") + 
  theme_minimal() + 
  labs(y = "Percent of school-aged females in school",
       x = "Percent with basic water access") +
  scale_color_manual(values = "#7463AC", name = "") 
```

## Poverty and water access 

```{r warning = FALSE}
# plot of poverty and water access
water.educ %>% 
  ggplot(aes(y = perc.basic2015water, x = perc.1dollar)) + 
  geom_point(size = 2, color = "#7463AC") + 
  theme_minimal() + 
  labs(y = "Percent with basic access to water",
       x = "Percent living on less than $1 per day")+
  scale_color_manual(values = "#7463AC", name = "") 
```

## Calculating Pearson's r correlation coefficient 

```{r}
# correlation between water access and female education
water.educ %>%
  summarize(cor.females.water = cor(x = perc.basic2015water,
                                    y = female.in.school, 
                                    use = "complete"))

```

## Water access and poverty in countries globally 

```{r warning=FALSE, echo=FALSE}
# explore plot of poverty and water
water.educ %>%
  ggplot(aes(x = perc.1dollar, y = perc.basic2015water)) + 
  geom_smooth(method = "lm", se = FALSE, aes(linetype = "Linear fit line"), color = "gray60") +
  geom_point(size = 2, color = "#7463AC") + 
  theme_minimal() + 
  labs(x = "Percent living on < $1 per day",
       y = "Percent with basic water access") +
  scale_linetype_manual(values = 1, name = "")
```

## Add correlation coefficient for poverty and water access 

```{r}
# correlations between water access, poverty, and female education
water.educ %>%
  summarize(cor.females.water = cor(x = perc.basic2015water,
                                    y = female.in.school,
                                    use = "complete"),
            cor.poverty.water = cor(x = perc.1dollar,
                                    y = perc.basic2015water,
                                    use = "complete"))

```

## You try it! Graph a relationship

Copy and modify the code from above to graph the relationship between the `female.in.school` variable and basic sanitation measured by the `perc.basic2015sani` variable.

```{r}
# relationship between female in school and sanitation
```

## You try it! Compute a correlation

Add the correlation between `female.in.school` and `perc.basic2015sani` to the existing code:

```{r}
# correlations between water access, poverty, and female education
water.educ %>%
  summarize(cor.females.water = cor(x = perc.basic2015water,
                                    y = female.in.school,
                                    use = "complete"),
            cor.poverty.water = cor(x = perc.1dollar,
                                    y = perc.basic2015water,
                                    use = "complete"))

```

## You try it! Interpret your results 


## NHST Step 2: Computing the test statistic 

```{r}
# test for correlation coefficient
cor.test(x = water.educ$perc.basic2015water, 
         y = water.educ$female.in.school)
```

## Checking normality using a histogram

```{r}
# check normality of female.in.school variable
water.educ %>%
  drop_na(female.in.school) %>%
  drop_na(perc.basic2015water) %>%
  ggplot(aes(x = female.in.school)) + 
  geom_histogram() + 
  theme_minimal() + 
  labs(x = "Percent of school-aged females in school",
       y = "Number of countries") 
```

## Checking normality for the water access variable

```{r}
# check normality of water access variable
water.educ %>%
  drop_na(female.in.school) %>%
  drop_na(perc.basic2015water) %>%
  ggplot(aes(x = perc.basic2015water)) + 
  geom_histogram() + 
  theme_minimal() + 
  labs(x = "Percent with basic water access",
       y = "Number of countries") 
```

## Checking the linearity assumption 

* Add a Loess curve to a scatterplot to check 

```{r}
# female education and water graph with linear fit line and Loess curve
water.educ %>%
  ggplot(aes(y = female.in.school, x = perc.basic2015water)) + 
  geom_point(aes(size = "Country"), color = "#7463AC") + 
  geom_smooth(aes(color = "Linear fit line"), method = "lm", se = FALSE) +
  geom_smooth(aes(color = "Loess curve"), se = FALSE) +
  theme_minimal() + 
  labs(y = "Percent of school-aged females in school",
       x = "Percent with basic access to water") +
  scale_color_manual(values = c("gray60", "deeppink"), name= "") +
  scale_size_manual(values = 2, name = "")
```
    
## Spearman's rho 

```{r message=FALSE, warning=FALSE}
# spearman correlation female education and water access
spear.fem.water <- cor.test(x = water.educ$perc.basic2015water, 
                            y = water.educ$female.in.school, 
                            method = "spearman")
spear.fem.water
```

## You try it! Test the correlation of sanitation with females in school

Step 1: Write the null and alternate hypotheses:

H0: There is no correlation between access to basic sanitation and percent of females in school. 

HA: There is a correlation between access to basic sanitation and percent of females in school.







