---
title: 'Applied Linear Modeling'
---


## Data source 

* The Foundation for AIDS Research (amFAR) includes distance to syringe exchange programs as one variable in the **Opioid & Health Indicators Database** (https://opioid.amfar.org/indicator/) 

* The `dist_ssp_amfar_ch9.csv` data set is county-level data that includes the distance to a syringe exchange program from counties in the US 

```{r}
# distance to syringe program data
dist.ssp <- read.csv(file = "dist_ssp_amfar_ch9.csv")

# summary
summary(object = dist.ssp)
```

## Codebook 

Based on the amFAR website, the variables have the following meanings:

* county: the county name 
* STATEABBREVIATION: the two-letter abbreviation for the state the county is in 
* dist_SSP: distance in miles to the nearest syringe services program 
* HIVprevalence: people age 13 and older living with diagnosed HIV per 100,000 
* opioid_RxRate: number of opioid prescriptions per 100 people 
* pctunins: percentage of the civilian noninstitutionalized population with no health insurance coverage 
* metro: county is non-metro, which includes open countryside, rural towns, or smaller cities with up to 49,999 people, or metro 

## A research question 

**How can uninsurance, metro or non-metro status, HIV prevalence, and number of opioid prescriptions predict or explain distance to the nearest syringe program at the county level?**

## But first, descriptives 

```{r}
# open the tidyverse
library(package = "tidyverse")

# checking distributions for continuous
dist.ssp %>%
  ggplot(aes(x = dist_SSP)) +
  geom_histogram(fill = "#7463AC", color = "white") +
  theme_minimal()
dist.ssp %>%
  ggplot(aes(x = HIVprevalence)) +
  geom_histogram(fill = "#7463AC", color = "white") +
  theme_minimal()
dist.ssp %>%
  ggplot(aes(x = opioid_RxRate)) +
  geom_histogram(fill = "#7463AC", color = "white") +
  theme_minimal()
dist.ssp %>%
  ggplot(aes(x = pctunins)) +
  geom_histogram(fill = "#7463AC", color = "white") +
  theme_minimal()

```

## Make the table 

```{r}
# make a table of descriptives
library(package = "tableone")

# dist_SSP and HIVprevalence are skewed
ssp.table <- CreateTableOne(data = dist.ssp,
                         vars = c('dist_SSP', 'HIVprevalence',
                                  'opioid_RxRate', 'pctunins',
                                  'metro'))
print(ssp.table, nonnormal = c("dist_SSP", "HIVprevalence"),
      showAllLevels = TRUE)
```

## Using R to compute the slope 

```{r}
# linear regression of distance to syringe program by percent uninsured
dist.by.unins <- lm(formula = dist_SSP ~ pctunins, 
                data = dist.ssp, na.action = na.exclude)
summary(dist.by.unins)
```

## Computing confidence intervals for the slope and intercept 

* Compute confidence intervals using the standard error of the slope using `confint()` function with the `dist.by.unins` linear regression model object 

```{r}
# confidence interval for regression parameters
ci.dist.by.unins <- confint(dist.by.unins)
ci.dist.by.unins
```

