---
title: "Week 1 workshop code"
output: html_document
--- 

```{r}
####################################
# Project name: 
# Project purpose: 
# Code author name: 
# Date last edited: 
# Location of data used: 
####################################
```

## Importing a data set into R 

Make sure this file and the data file are saved in the same folder before you begin. If they are not, you will need to change the file location of the data file for data importing.

```{r}
# read the GSS 2016 data 
gss.2016 <- read.csv(file = "legal_weed_age_GSS2016_ch1.csv")
```

## Examine the contents of the data file 

```{r}
# examine the contents of the file
summary(object = gss.2016)
```

## Another way to import csv data 

```{r}
# open the data.table package 
library(package = "data.table")
```

## Using fread to import data

```{r}
# open the GSS data with fread
gss.2016 <- fread(input = "legal_weed_age_GSS2016_ch1.csv")
```

## Try the summary function again to see if it looks better 

```{r}
# examine the contents of the file
summary(object = gss.2016)
```

## Check data types using the class function 

```{r}
# what data type is grass variable
class(x = gss.2016$grass)

# what data type is age variable
class(x = gss.2016$age)
```

## Fixing the data types 

```{r}
# open the tidyverse 
library(package = "tidyverse")
```

## Our first data cleaning task! 

```{r}
# change the data types for grass and age 
gss.2016.clean <- gss.2016 %>% 
    mutate(grass = as.factor(grass)) %>%
    mutate(age = as.numeric(age))
```

## Checking our work 

```{r}
# check the types
class(x = gss.2016.clean$grass)
class(x = gss.2016.clean$age)

# use summary to check the new data 
summary(object = gss.2016.clean)
```

## Coding missing values 

```{r}
# recode don't know and inapplicable to NA
gss.2016.clean <- gss.2016 %>% 
    mutate(grass = as.factor(grass)) %>%
    mutate(age = as.numeric(age)) %>% 
    mutate(grass = na_if(x = grass, y = "DK")) %>% 
    mutate(grass = na_if(x = grass, y = "IAP"))

```

## Check our work 

```{r}
# check for NA
summary(object = gss.2016.clean)
```

## Get rid of unused categories of a factor 

```{r}
# recode don't know and inapplicable to NA
gss.2016.clean <- gss.2016 %>% 
    mutate(grass = as.factor(grass)) %>%
    mutate(age = as.numeric(age)) %>% 
    mutate(grass = na_if(x = grass, y = "DK")) %>% 
    mutate(grass = na_if(x = grass, y = "IAP")) %>% 
    mutate(grass = droplevels(x = grass))

```

## Check our work 

```{r}
# check for dropped levels
summary(object = gss.2016.clean)
```

## Let's recode age to be categorical!

```{r}
# recode don't know and inapplicable to NA
gss.2016.clean <- gss.2016 %>% 
    mutate(grass = as.factor(grass)) %>%
    mutate(age = as.numeric(age)) %>% 
    mutate(grass = na_if(x = grass, y = "DK")) %>% 
    mutate(grass = na_if(x = grass, y = "IAP")) %>% 
    mutate(grass = droplevels(x = grass)) %>%
    mutate(age.cat = cut(x = age,
                         breaks = c(18, 29, 59, 74, 88),
                         labels = c("18 - 29", "30 - 59", "60 - 74", "75+")))

```

## Let's rename the legalization categories

```{r}
# recode don't know and inapplicable to NA
gss.2016.clean <- gss.2016 %>% 
    mutate(grass = as.factor(grass)) %>%
    mutate(age = as.numeric(age)) %>% 
    mutate(grass = na_if(x = grass, y = "DK")) %>% 
    mutate(grass = na_if(x = grass, y = "IAP")) %>% 
    mutate(grass = droplevels(x = grass)) %>% 
    mutate(grass = recode_factor(grass, 
                          `LEGAL` = "Yes",
                          `NOT LEGAL` = "No")) %>%
    mutate(age.cat = cut(x = age,
                         breaks = c(18, 29, 59, 74, 88),
                         labels = c("18 - 29", "30 - 59", "60 - 74", "75+")))

```

## Check our work 

```{r}
# check for age.cat variable 
summary(object = gss.2016.clean)
```

## Let's make a graph! 

```{r}
# graph support for legalization 
gss.2016.clean %>%
    ggplot(aes(x = grass)) +
    geom_bar()
```

## Let's get rid of the NA in the graph

```{r}
# graph support for legalization 
gss.2016.clean %>%
    drop_na(grass) %>%
    ggplot(aes(x = grass)) +
    geom_bar()
```

## Let's add some color

```{r}
# graph support for legalization 
gss.2016.clean %>%
    drop_na(grass) %>%
    ggplot(aes(x = grass, fill = grass)) +
    geom_bar()
```

## How about green instead? ...and we do not need the legend

```{r}
# graph support for legalization 
gss.2016.clean %>%
    drop_na(grass) %>%
    ggplot(aes(x = grass, fill = grass)) + 
    geom_bar() +
    scale_fill_manual(values = c("#78A678", "#7463AC"), 
                    guide = FALSE)
```

## Let's add better labels to the axes 

```{r}
# graph support for legalization 
gss.2016.clean %>%
    drop_na(grass) %>%
    ggplot(aes(x = grass, fill = grass)) + 
    geom_bar() +
    scale_fill_manual(values = c("#78A678", "#7463AC"), 
                    guide = FALSE) +
    labs(x = "Should marijuana be legal?",
         y = "Number of responses")
```

## Get rid of the gray background 

```{r}
# graph support for legalization 
gss.2016.clean %>%
    drop_na(grass) %>%
    ggplot(aes(x = grass, fill = grass)) + 
    geom_bar() +
    scale_fill_manual(values = c("#78A678", "#7463AC"), 
                    guide = FALSE) +
    labs(x = "Should marijuana be legal?",
         y = "Number of responses") +
    theme_minimal()
```

## Add age to the graph

```{r}
# graph support for legalization 
gss.2016.clean %>%
    drop_na(grass) %>%
    drop_na(age.cat) %>%
    ggplot(aes(x = age.cat, fill = grass)) + 
    geom_bar(position = "dodge") +
    scale_fill_manual(values = c("#78A678", "#7463AC"), 
                    guide = FALSE) +
    labs(x = "Should marijuana be legal?",
         y = "Number of responses") +
    theme_minimal()
```
